Constraint result to the self-interacting dark matter model without dark energy in cosmology

### 2025-04-09 article/article.tex
seperate the data caculating and figure plotting

   The standard Λ-Cold Dark Matter (ΛCDM) model has been successful in 
   explaining the accelerated expansion of the Universe (\cite{Riess_1998}; \cite{Perlmutter_1999}), 
   from the cosmic microwave background (CMB) anisotropies (\cite{Bennett_1996}) to 
   the measurement of baryon acoustic oscillations (BAO; \cite{Eisenstein_2005}). However, there are also many 
   alternative models that have been proposed to explain, such as the dynamical
   dark energy models and the modified gravity models.

   It has been shown that dark matter self-interaction could distribute 
   to the accelerated expansion without dark energy component. According
   to the Boltzmann formalism, the disequilibrium between dark matter particle
   creation and annihilation processes would creat an effective source term 
   making negative pressure just like the dark energy. \cite{Basilakos_2009}
   investigated the analytical solutions for the simple interacting dark matter (IDM) 
   model, and find that the effective annihilation term is much smaller than the
   results given by the general methods. The gravitational matter creation
   model is mathematically equivalent to one case of the IDM models.

@@

   We assume that the total density of the cosmic fluid obeys
   the collisional Boltzmann equation
   \begin{equation}
      \dot{\rho}+3H\rho+\kappa\rho^2-2\Psi=0,
   \end{equation}
   where $\rho$ is the total energy-density of the cosmic fluid,
   containing dark matter, baryons, and any type of exotic energy,
   $\Psi$ is the rate of creation of DM particle pairs, and the
   annihilation parameter $\kappa(\geq0)$ is given by:
   \begin{equation}
      \kappa=\frac{\langle\sigma u\rangle}{M_x},\label{eq:2}
   \end{equation}
   where $\sigma$ is the cross-section for annihilation, $u$ is
   the mean particle velocity, and $M_x$ is the mass of the DM
   particle.
   Compared to the usual fluid equation, the effective pressure term
   is \begin{equation}
      P=\frac{\kappa\rho^2-\Psi}{3H}.
   \end{equation}
   When $\kappa\rho^2-\Psi<0,$ what means that the IDM particle
   creation term is larger than the annihilation item, IDM may serve
   as a negative pressure source in the global dynamics of the Universe,
   like the role of Dark Energy in the general cosmological models.

@@

   If $\kappa=0$, the global density evolution can be written as
   \begin{equation}
      \rho(a)=\mathcal{D}a^{-3}+C_1a^{n},
   \end{equation}
   where $\mathcal{D}=C_2-C-1$. The conditions in which the current 
   model acts as a quintessence cosmology are given by $\mathcal{D}>0,
   C_1>0,$ and $w_{\text{IDM}}=-1-n/3$. This solution is mathematically
   equivalent to that of the gravitational matter creation model of().
   The Hubble flow is now given by
   \begin{equation}
      \left(\frac{H}{H_0}\right)^2=\Omega_{2,0}a^{-3}+\Omega_{1,0}a^{n},
   \end{equation}
   where $\Omega_{2,0}=8\pi G\mathcal{D}/3H_0^2$ and 
   $\Omega_{1,0}=8\pi GC_1/3H_0^2$, respectively.(\cite{Basilakos_2009})

@@

   To constrain the relevant IDM models (\cite{Basilakos_2009}), we use the newly revised
   observational $H(z)$ data (OHD)(\cite{PhysRevD.71.123001}; 
   \cite{Daniel.Stern_2010}; \cite{M.Moresco_2012}; \cite{Zhang_2014};
   \cite{Moresco_2016}; \cite{10.1093/mnras/stx301}; \cite{10.1093/mnrasl/slv037}; 
   \cite{Borghi_2022}; \cite{Jiao_2023}), the Pantheon+ set of 1701 SNe Ia 
   (\cite{Scolnic_2022}), the quasar data from \cite{Lusso_2020}, 
   the BAO data from SDSS and DESI 2024.

@@

   \begin{table}[htbp]
      \caption{The OHD dataset}
      \centering
      \begin{tabular}{llll}
         \hline\hline
         $z$ & $H(z)$ & Reference \\
         \hline     
         0.07 & 69$\pm$19.6 & \cite{Zhang_2014} \\
         0.09 & 69$\pm$12 & \cite{PhysRevD.71.123001} \\
         0.12 & 68.6$\pm$26.2 & \cite{Zhang_2014} \\
         0.17 & 83$\pm$8 & \cite{PhysRevD.71.123001} \\
         0.179 & 75$\pm$4 & \cite{M.Moresco_2012} \\
         0.199 & 75$\pm$5 & \cite{M.Moresco_2012} \\
         0.2 & 72.9$\pm$29.6 & \cite{Zhang_2014} \\
         0.27 & 77$\pm$14 & \cite{PhysRevD.71.123001} \\
         0.28 & 88.8$\pm$36.6 & \cite{Zhang_2014} \\
         0.352 & 83$\pm$14 & \cite{M.Moresco_2012} \\
         0.3802 & 83$\pm$13.5 & \cite{Moresco_2016}  \\
         0.4 & 95$\pm$17 & \cite{PhysRevD.71.123001} \\
         0.4004 & 77$\pm$10.2 & \cite{Moresco_2016} \\
         0.4247 & 87.1$\pm$11.2 & \cite{Moresco_2016} \\
         0.4497 & 92.8$\pm$12.9 & \cite{Moresco_2016} \\
         0.47 & 89$\pm$34 & \cite{10.1093/mnras/stx301} \\
         0.4783 & 80.9$\pm$9 & \cite{Moresco_2016} \\
         0.48 & 97$\pm$62 & \cite{Daniel.Stern_2010} \\
         0.593 & 104$\pm$13 & \cite{M.Moresco_2012} \\
         0.68 & 92$\pm$8 & \cite{M.Moresco_2012} \\
         0.75 & 98.8$\pm$33.6 & \cite{Borghi_2022} \\
         0.781 & 105$\pm$12 & \cite{M.Moresco_2012} \\
         0.8 & 113.1$\pm$15.1 & \cite{Jiao_2023} \\
         0.875 & 125$\pm$17 & \cite{M.Moresco_2012} \\
         0.88 & 90$\pm$40 & \cite{Daniel.Stern_2010} \\
         0.9 & 117$\pm$23 & \cite{PhysRevD.71.123001} \\
         1.037 & 154$\pm$20 & \cite{M.Moresco_2012} \\
         1.3 & 168$\pm$17 & \cite{PhysRevD.71.123001} \\
         1.363 & 160$\pm$33.6 & \cite{10.1093/mnrasl/slv037} \\
         1.43 & 177$\pm$18 & \cite{PhysRevD.71.123001} \\
         1.53 & 140$\pm$14 & \cite{PhysRevD.71.123001} \\
         1.75 & 202$\pm$40 & \cite{PhysRevD.71.123001} \\
         1.965 & 186.5$\pm$50.4 & \cite{10.1093/mnrasl/slv037} \\
         \hline    
      \end{tabular}
      \label{tab:1}
   \end{table}

@@

   SNe Ia have long been used as "standard candles" to give a direct
   measurement of their luminosity distance, and provides strong constraints
   on cosmological parameters. We use the latest Pantheon+ data set of 1701 
   SNe Ia samples(\cite{Scolnic_2022}), which covers the redshift range [0, 2.26].

   We use the fiducial SN Ia magnitude ($M_b$) determined from SH0ES 2021 Cepheid 
   host distances(\cite{Riess_2022}), which gives the $\mu_{\text{data}}$ and we 
   give the $\chi^2$ as 
   \begin{equation}
      \chi_{\text{SNe}}^2=\Delta^{\text{T}}C^{-1}\Delta,
   \end{equation}
   where $\Delta=(\mu_{\text{th}}-\mu_{\text{data}})$ and $C^{-1}$ is the inverse of the
   covariance matrix of the SNe Ia data, the distance modulus is $\mu=5\log_{10}(d_L/\text{Mpc})+25$, and the
   luminosity distance $d_L$ can be given as a function of redshift $z$
   \begin{equation}
      d_L=(1+z)\int_0^z\frac{c\mathrm{d}z'}{H(z')}.\label{eq:15}
   \end{equation}
   To eliminate the advanced restriction to $H_0$ from $M_b$, we adopt
   the likelihood function as
   \begin{equation}
      \widetilde{\chi}_{\text{SNe}}^2=\chi_{\text{SNe}}^2-\frac{B^2}{C}+\ln\left(\frac{C}{2\pi}\right),
   \end{equation}
   where $B=\Delta^{\text{T}}C^{-1}$ and $C$ is the sum of $C^{-1}$.
   Apparently these two functions give the same constraints in $\Omega_{2,0}$ and
   $\log_{10}(\kappa C_1)$.

@@

   The Baryon acoustic oscillation method (BAO) provides a key cosmological probe
   sensitive to the cosmic expansion history with well-controlled systematics.
   We use two BAO data sets from the SDSS(\cite{PhysRevD.103.083533}) and DESI 2024(\cite{desicollaboration2024desi2024vicosmological}),
   which are given at Table \ref{tab:2} and Table \ref{tab:3}, respectively.
   The redshift is up to 2.33 both in the SDSS and the DESI 2024 dataset.

@@

   \begin{table}[htbp]
      \caption{The BAO-only dataset from SDSS}
      \centering
      \begin{tabular}{llll}
         \hline\hline
         $z_{\text{eff}}$ & $D_{\text{M}}/r_{\text{d}}$ & 
         $D_{\text{H}}/r_{\text{d}}$ & $D_{\text{V}}/r_{\text{d}}$ \\
         \hline
         0.15 & & & 4.47$\pm$0.17 \\
         0.38 & 10.23$\pm$0.17 & 25$\pm$0.76 & \\
         0.51 & 13.36$\pm$0.21 & 22.33$\pm$0.58 & \\
         0.7 & 17.86$\pm$0.33 & 19.33$\pm$0.53 & \\
         0.85 & & & $18.33_{-0.62}^{+0.57}$ \\
         1.48 & 30.69$\pm$0.8 & 13.26$\pm$0.55 & \\
         2.33 & 37.6$\pm$1.9 & 8.93$\pm$0.28 & \\
         2.33 & 37.3$\pm$1.7 & 9.08$\pm$0.34 & \\
         \hline
      \end{tabular}
      \label{tab:2}
   \end{table}

   \begin{table}[htbp]
      \caption{The BAO dataset from DESI 2024}
      \centering
      \begin{tabular}{llll}
         \hline\hline
         $z_{\text{eff}}$ & $D_{\text{M}}/r_{\text{d}}$ & 
         $D_{\text{H}}/r_{\text{d}}$ & $D_{\text{V}}/r_{\text{d}}$ \\
         \hline
         0.295 & & & 7.93$\pm$0.15 \\
         0.51 & 13.62$\pm$0.25 & 20.98$\pm$0.61 & \\
         0.706 & 16.85$\pm$0.32 & 20.08$\pm$0.6 & \\
         0.93 & 21.71$\pm$0.28 & 17.88$\pm$0.35 & \\
         1.317 & 27.79$\pm$0.69 & 13.82$\pm$0.42 & \\
         1.491 & & & 26.07$\pm$0.67 \\
         2.33 & 39.71$\pm$0.94 & 8.52$\pm$0.17 & \\
         \hline
      \end{tabular}
      \label{tab:3}
   \end{table}

@@

   We use the Markov chain Monte Carlo (MCMC) method based 
   on the opening package \texttt{emcee} to give a global constraints
   to the free parameters $\Omega_{2,0}$ and $\log_{10}(\kappa C_1)$ in 
   Model 1 and $n$ in Model 2.
   Besides, we also add the parameter $H_0$ or $h$ to give the constraints
   to the dark matter particle mass $M_x$.

   The prior range set for free parameters are given at Table \ref{tab:4}

   \begin{table}[htbp]
      \caption{Parameters and priors used in analysis}
      \centering
      \begin{tabular}{lll}
         \hline\hline
         parameter & initial & prior \\
         \hline
         $\Omega_{2,0}$ & 0.3 & $\mathcal{U}[0.0,1.0]$ \\
         $n$ & 0 & $\mathcal{U}[-10,10]$ \\
         $\log_{10}(\kappa C_1/$Gyr${}^{-1})$ & --- & $\mathcal{U}[-10,0]$ \\
         $H_0$[km/s/Mpc] & 70 & $\mathcal{U}[60,80]$ \\
         \hline
         $\beta_0$ & 6 & $\mathcal{U}[-15,15]$ \\
         $\beta_1$ & 1 & $\mathcal{U}[-10,10]$ \\
         $\gamma_0$ & 0.6 & $\mathcal{U}[0.0,1.0]$ \\
         $\gamma_1$ & 0 & $\mathcal{U}[-1.0,1.0]$ \\         
         $\delta$ & 0.2 & $\mathcal{U}[0.0,1.0]$ \\
         \hline
         $r_{\text{d}}h$[Mpc] & 100 & $\mathcal{U}[50,150]$ \\
         \hline
      \end{tabular}
      \label{tab:4}
   \end{table}

@@

   The Eq.(\ref{eq:8}) convergant to the flat ΛCDM model as
   $\log_{10}(\kappa C_1)\to-\infty$, therefore the constraints
   could only give the upper limit of what, and give the lower limit
   of $M_x$ according to Eq.(\ref{eq:9}), respectively.

   We use the 97.7\% and 2.3\% quantiles (about $2\sigma$) to determine the upper and lower
   limits of the parameters, and the results are shown in Table \ref{tab:5}.

   \begin{figure}[htbp]
      \centering
      \subfigure{
         \includegraphics[width=0.22\textwidth]{ohd_1.pdf}
      }
      \subfigure{
         \includegraphics[width=0.22\textwidth]{ohd_2.pdf}
      }
      \caption{The constraint result from the observational
      $H(z)$ data}
   \end{figure}

   \begin{figure}[htbp]
      \centering
      \subfigure{
         \includegraphics[width=0.22\textwidth]{sne_1.pdf}
      }
      \subfigure{
         \includegraphics[width=0.22\textwidth]{sne_2.pdf}
      }
      \caption{The constraint result from the supernovae
      Ia data}
   \end{figure}

   \begin{figure}[htbp]
      \centering
      \subfigure{
         \includegraphics[width=0.22\textwidth]{sne_qso_1.pdf}
      }
      \subfigure{
         \includegraphics[width=0.22\textwidth]{sne_qso_2.pdf}
      }
      \caption{The constraint result from the SNe Ia + QSO data}
   \end{figure}

   We would find that the SNe Ia and QSO combined result is most
   similar to SNe Ia result, what may refer that the QSO cannot
   give an effective constraint. It may be due to the data quality
   at the high redshift.

   \begin{figure}[htbp]
      \centering
      \subfigure{
         \includegraphics[width=0.22\textwidth]{bao_1.pdf}
      }
      \subfigure{
         \includegraphics[width=0.22\textwidth]{bao_2.pdf}
      }
      \caption{The constraint result from the BAO data}
   \end{figure}

   \begin{figure}[htbp]
      \centering
      \subfigure{
         \includegraphics[width=0.22\textwidth]{ohd_sne_1.pdf}
      }
      \subfigure{
         \includegraphics[width=0.22\textwidth]{ohd_sne_2.pdf}
      }
      \caption{The constraint result from the OHD + SNe Ia data}
   \end{figure}

   \begin{figure}[htbp]
      \centering
      \subfigure{
         \includegraphics[width=0.22\textwidth]{ohd_sne_bao_1.pdf}
      }
      \subfigure{
         \includegraphics[width=0.22\textwidth]{ohd_sne_bao_2.pdf}
      }
      \caption{The constraint result from the OHD + SNe Ia + BAO data}
   \end{figure}

   \begin{table*}[htbp]
      \caption{The constraint results of Model 1}
      \centering
      \begin{tabular}{lcccccc}
         \hline\hline
         Model 1 & OHD & SNe Ia & SNe Ia + QSO & BAO &
          OHD + SNe Ia & OHD + SNe Ia + BAO \\
         \hline
         $\Omega_{2,0}$ & $0.321_{-0.067}^{+0.055}$ & $0.360_{-0.020}^{+0.020}$
          & $0.359_{-0.020}^{+0.020}$ & $0.301_{-0.013}^{+0.013}$
          & $0.330_{-0.016}^{+0.016}$ & $0.313_{-0.010}^{+0.010}$ \\
         $\log_{10}(\kappa C_1/$Gyr${}^{-1})$ & $<-2.17$ & $<-2.51$ & $<-2.57$
          & $<-3.10$ & $<-2.88$ & $<-3.39$ \\
         $\log_{10}(M_x/$GeV) & $>-9.77$ & $>-9.42$ & $>-9.40$
          & $>-8.84$ & $>-9.03$ & $>-8.52$ \\
         \hline
      \end{tabular}
      \label{tab:5}
   \end{table*}

@@

   \begin{figure}[htbp]
      \centering
      \subfigure{
         \includegraphics[width=0.22\textwidth]{ohd_widm_1.pdf}
      }
      \subfigure{
         \includegraphics[width=0.22\textwidth]{ohd_widm_2.pdf}
      }
      \caption{The constraint result from the observational
      $H(z)$ data}
   \end{figure}

   \begin{figure}[htbp]
      \centering
      \subfigure{
         \includegraphics[width=0.22\textwidth]{sne_widm_1.pdf}
      }
      \subfigure{
         \includegraphics[width=0.22\textwidth]{sne_widm_2.pdf}
      }
      \caption{The constraint result from the supernovae
      Ia data}
   \end{figure}

   \begin{figure}[htbp]
      \centering
      \subfigure{
         \includegraphics[width=0.22\textwidth]{sne_qso_widm_1.pdf}
      }
      \subfigure{
         \includegraphics[width=0.22\textwidth]{sne_qso_widm_2.pdf}
      }
      \caption{The constraint result from the SNe Ia + QSO data}
   \end{figure}

>>>>

   The standard Λ-Cold Dark Matter (ΛCDM) model has been successful in 
   explaining the accelerated expansion of the Universe (\cite{Riess_1998,Perlmutter_1999}), 
   from the cosmic microwave background (CMB) anisotropies (\cite{Bennett_1996}) to 
   the measurement of baryon acoustic oscillations (BAO; \cite{Eisenstein_2005}). 
   However, there are also many alternative models that have been proposed to describe 
   the phenomenon, such as the canonical ``quintessence" scalar field theories 
   (\cite{PhysRevD.37.3406, WETTERICH1988668,PhysRevLett.80.1582}),
   non-canonical ``k-essence" theories (\cite{PhysRevLett.85.4438, PhysRevD.63.103510})
   and the modified gravity models (see, e.g., \cite{CLIFTON20121} for a review).

   It has been shown that dark matter self-interaction could contribute 
   to the accelerated expansion without dark energy component (\cite{PhysRevD.64.063501, Balakin_2003}). 
   According to the Boltzmann formalism, the disequilibrium between dark matter particle
   creation and annihilation processes would creat an effective source term 
   making negative pressure just like the dark energy. \cite{Basilakos_2009}
   investigated the analytical solutions for the simple interacting dark matter (IDM) 
   model, and find that the effective annihilation term is much smaller than the
   results given by the general methods. The gravitational matter creation model (\cite{Lima_2008})
   is mathematically equivalent to one case of the IDM models (\cite{Basilakos_2009}).

@@

   We assume that the total density of the cosmic fluid obeys
   the collisional Boltzmann equation (\cite{Kolb:1990vq})
   \begin{equation}
      \dot{\rho}+3H\rho+\kappa\rho^2-2\Psi=0,
   \end{equation}
   where $\rho$ is the total energy-density of the cosmic fluid,
   containing dark matter, baryons, and any type of exotic energy,
   $\Psi$ is the rate of creation of DM particle pairs, and the
   annihilation parameter $\kappa(\geq0)$ is given by:
   \begin{equation}
      \kappa=\frac{\langle\sigma u\rangle}{M_x},\label{eq:2}
   \end{equation}
   where $\sigma$ is the cross-section for annihilation, $u$ is
   the mean particle velocity, and $M_x$ is the mass of the DM
   particle.
   Compared to the usual fluid equation, the effective pressure term
   is \begin{equation}
      P=\frac{\kappa\rho^2-\Psi}{3H}.
   \end{equation}
   When $\kappa\rho^2-\Psi<0,$ what means that the IDM particle
   creation term is larger than the annihilation item, IDM may serve
   as a negative pressure source in the global dynamics of the Universe,
   like the role of Dark Energy in the general cosmological models.

@@

   If $\kappa=0$, the global density evolution can be written as
   \begin{equation}
      \rho(a)=\mathcal{D}a^{-3}+C_1a^{n},
   \end{equation}
   where $\mathcal{D}=C_2-C_1$. The conditions in which the current 
   model acts as a quintessence cosmology are given by $\mathcal{D}>0,
   C_1>0,$ and $w_{\text{IDM}}=-1-n/3$. This solution is mathematically
   equivalent to that of the gravitational matter creation model of \cite{Lima_2008}.
   The Hubble flow is now given by
   \begin{equation}
      \left(\frac{H}{H_0}\right)^2=\Omega_{2,0}a^{-3}+\Omega_{1,0}a^{n},
   \end{equation}
   where $\Omega_{2,0}=8\pi G\mathcal{D}/3H_0^2$ and 
   $\Omega_{1,0}=8\pi GC_1/3H_0^2$, respectively.(\cite{Basilakos_2009})

@@

   To constrain the relevant IDM models (\cite{Basilakos_2009}), we use the newly revised
   observational $H(z)$ data (OHD)(\cite{PhysRevD.71.123001, Daniel.Stern_2010, 
   M.Moresco_2012, Zhang_2014, Moresco_2016, 10.1093/mnras/stx301, 10.1093/mnrasl/slv037, 
   Borghi_2022, Jiao_2023}), the Pantheon+ set of 1701 SNe Ia 
   (\cite{Scolnic_2022}), the quasar data from \cite{Lusso_2020}, 
   the BAO data from SDSS (\cite{PhysRevD.103.083533}) and DESI 2024 
   (\cite{desicollaboration2024desi2024vicosmological}).

@@

   \begin{table}[htbp]
      \centering
      \begin{tabular}{llll}
         \hline\hline
         $z$ & $H(z)$ & Reference \\
         \hline     
         0.07 & 69$\pm$19.6 & \cite{Zhang_2014} \\
         0.09 & 69$\pm$12 & \cite{PhysRevD.71.123001} \\
         0.12 & 68.6$\pm$26.2 & \cite{Zhang_2014} \\
         0.17 & 83$\pm$8 & \cite{PhysRevD.71.123001} \\
         0.179 & 75$\pm$4 & \cite{M.Moresco_2012} \\
         0.199 & 75$\pm$5 & \cite{M.Moresco_2012} \\
         0.2 & 72.9$\pm$29.6 & \cite{Zhang_2014} \\
         0.27 & 77$\pm$14 & \cite{PhysRevD.71.123001} \\
         0.28 & 88.8$\pm$36.6 & \cite{Zhang_2014} \\
         0.352 & 83$\pm$14 & \cite{M.Moresco_2012} \\
         0.3802 & 83$\pm$13.5 & \cite{Moresco_2016}  \\
         0.4 & 95$\pm$17 & \cite{PhysRevD.71.123001} \\
         0.4004 & 77$\pm$10.2 & \cite{Moresco_2016} \\
         0.4247 & 87.1$\pm$11.2 & \cite{Moresco_2016} \\
         0.4497 & 92.8$\pm$12.9 & \cite{Moresco_2016} \\
         0.47 & 89$\pm$34 & \cite{10.1093/mnras/stx301} \\
         0.4783 & 80.9$\pm$9 & \cite{Moresco_2016} \\
         0.48 & 97$\pm$62 & \cite{Daniel.Stern_2010} \\
         0.593 & 104$\pm$13 & \cite{M.Moresco_2012} \\
         0.68 & 92$\pm$8 & \cite{M.Moresco_2012} \\
         0.75 & 98.8$\pm$33.6 & \cite{Borghi_2022} \\
         0.781 & 105$\pm$12 & \cite{M.Moresco_2012} \\
         0.8 & 113.1$\pm$15.1 & \cite{Jiao_2023} \\
         0.875 & 125$\pm$17 & \cite{M.Moresco_2012} \\
         0.88 & 90$\pm$40 & \cite{Daniel.Stern_2010} \\
         0.9 & 117$\pm$23 & \cite{PhysRevD.71.123001} \\
         1.037 & 154$\pm$20 & \cite{M.Moresco_2012} \\
         1.3 & 168$\pm$17 & \cite{PhysRevD.71.123001} \\
         1.363 & 160$\pm$33.6 & \cite{10.1093/mnrasl/slv037} \\
         1.43 & 177$\pm$18 & \cite{PhysRevD.71.123001} \\
         1.53 & 140$\pm$14 & \cite{PhysRevD.71.123001} \\
         1.75 & 202$\pm$40 & \cite{PhysRevD.71.123001} \\
         1.965 & 186.5$\pm$50.4 & \cite{10.1093/mnrasl/slv037} \\
         \hline    
      \end{tabular}
      \caption{The OHD dataset from different paper.}
      \label{tab:1}
   \end{table}

@@

   SNe Ia have long been used as ``standard candles" to give a direct
   measurement of their luminosity distance, and provides strong constraints
   on cosmological parameters. We use the latest Pantheon+ data set of 1701 
   SNe Ia samples (\cite{Scolnic_2022}), which covers the redshift range [0, 2.26].

   We use the fiducial SNe Ia magnitude ($M_b$) determined from SH0ES 2021 Cepheid 
   host distances (\cite{Riess_2022}), which gives the $\mu_{\text{data}}$ and we 
   give the $\chi^2$ as 
   \begin{equation}
      \chi_{\text{SNe}}^2=\Delta^{\text{T}}C^{-1}\Delta,
   \end{equation}
   where $\Delta=(\mu_{\text{th}}-\mu_{\text{data}})$ and $C^{-1}$ is the inverse of the
   covariance matrix of the SNe Ia data, the distance modulus is $\mu=5\log_{10}(d_L/\text{Mpc})+25$, and the
   luminosity distance $d_L$ can be given as a function of redshift $z$
   \begin{equation}
      d_L=(1+z)\int_0^z\frac{c\mathrm{d}z'}{H(z')}.\label{eq:15}
   \end{equation}
   To eliminate the advanced restriction to $H_0$ from $M_b$, we adopt
   the likelihood function as
   \begin{equation}
      \widetilde{\chi}_{\text{SNe}}^2=\chi_{\text{SNe}}^2-\frac{B^2}{C}+\ln\left(\frac{C}{2\pi}\right),
   \end{equation}
   where $B=\Delta^{\text{T}}C^{-1}$ and $C$ is the sum of $C^{-1}$.
   Apparently, these two functions give the same constraints in $\Omega_{2,0}$ and
   $\log_{10}(\kappa C_1)$.

@@

   The Baryon acoustic oscillation (BAO) method provides a key cosmological probe
   sensitive to the cosmic expansion history with well-controlled systematics.
   We use two BAO datasets from the SDSS (\cite{PhysRevD.103.083533}) and DESI 2024 (\cite{desicollaboration2024desi2024vicosmological}),
   which are given at Table \ref{tab:2} and Table \ref{tab:3}, respectively.
   The redshift is up to 2.33 both in the SDSS and the DESI 2024 dataset.

@@

   \begin{table}[htbp]
      \centering
      \begin{tabular}{llll}
         \hline\hline
         $z_{\text{eff}}$ & $D_{\text{M}}/r_{\text{d}}$ & 
         $D_{\text{H}}/r_{\text{d}}$ & $D_{\text{V}}/r_{\text{d}}$ \\
         \hline
         0.15 & & & 4.47$\pm$0.17 \\
         0.38 & 10.23$\pm$0.17 & 25$\pm$0.76 & \\
         0.51 & 13.36$\pm$0.21 & 22.33$\pm$0.58 & \\
         0.7 & 17.86$\pm$0.33 & 19.33$\pm$0.53 & \\
         0.85 & & & $18.33_{-0.62}^{+0.57}$ \\
         1.48 & 30.69$\pm$0.8 & 13.26$\pm$0.55 & \\
         2.33 & 37.6$\pm$1.9 & 8.93$\pm$0.28 & \\
         2.33 & 37.3$\pm$1.7 & 9.08$\pm$0.34 & \\
         \hline
      \end{tabular}
      \caption{The BAO-only dataset from SDSS. The $\sigma$ of asymmetric 
      uncertainties would be calculated on average.}
      \label{tab:2}
   \end{table}

   \begin{table}[htbp]
      \centering
      \begin{tabular}{llll}
         \hline\hline
         $z_{\text{eff}}$ & $D_{\text{M}}/r_{\text{d}}$ & 
         $D_{\text{H}}/r_{\text{d}}$ & $D_{\text{V}}/r_{\text{d}}$ \\
         \hline
         0.295 & & & 7.93$\pm$0.15 \\
         0.51 & 13.62$\pm$0.25 & 20.98$\pm$0.61 & \\
         0.706 & 16.85$\pm$0.32 & 20.08$\pm$0.6 & \\
         0.93 & 21.71$\pm$0.28 & 17.88$\pm$0.35 & \\
         1.317 & 27.79$\pm$0.69 & 13.82$\pm$0.42 & \\
         1.491 & & & 26.07$\pm$0.67 \\
         2.33 & 39.71$\pm$0.94 & 8.52$\pm$0.17 & \\
         \hline
      \end{tabular}
      \caption{The BAO dataset from DESI 2024.}
      \label{tab:3}
   \end{table}

@@

   We use the Markov chain Monte Carlo (MCMC) method based 
   on the opening package \texttt{emcee} to give a global constraints
   to the free parameters $\Omega_{2,0}$ and $\log_{10}(\kappa C_1)$ in 
   Model 1 and $n$ in Model 2.
   Besides, we also add the parameter $H_0$ or $h$ to give the constraints
   to the dark matter particle mass $M_x$.
   The prior range set for free parameters are given at Table \ref{tab:4}

   \begin{table}[htbp]
      \centering
      \begin{tabular}{lll}
         \hline\hline
         parameter & initial & prior \\
         \hline
         $\Omega_{2,0}$ & 0.3 & $\mathcal{U}[0.0,1.0]$ \\
         $n$ & 0 & $\mathcal{U}[-10,10]$ \\
         $\log_{10}(\kappa C_1/$Gyr${}^{-1})$ & --- & $\mathcal{U}[-10,0]$ \\
         $H_0$[km/s/Mpc] & 70 & $\mathcal{U}[60,80]$ \\
         \hline
         $\beta_0$ & 6 & $\mathcal{U}[-15,15]$ \\
         $\beta_1$ & 1 & $\mathcal{U}[-10,10]$ \\
         $\gamma_0$ & 0.6 & $\mathcal{U}[0.0,1.0]$ \\
         $\gamma_1$ & 0 & $\mathcal{U}[-1.0,1.0]$ \\         
         $\delta$ & 0.2 & $\mathcal{U}[0.0,1.0]$ \\
         \hline
         $r_{\text{d}}h$[Mpc] & 100 & $\mathcal{U}[50,150]$ \\
         \hline
      \end{tabular}
      \caption{Parameters and priors used in analysis. 
      All of the priors are flat in the ranges given. 
      The $\log_{10}(\kappa C_1)$ does not have a predicted value, 
      and the initial value of each chain is selected randomly uniformly
      within the prior range.}
      \label{tab:4}
   \end{table}

   For the BAO dataset, we use $h \simeq 0.7$ as the default value to give constraints.

@@

   The Eq.(\ref{eq:8}) convergant to the flat ΛCDM model as
   $\log_{10}(\kappa C_1)\to-\infty$, thus the constraints
   could only give the upper limit of what, and give the lower limit
   of $M_x$ according to Eq.(\ref{eq:9}), respectively.

   We use the 97.7\% and 2.3\% quantiles (about $2\sigma$) to determine the upper and lower
   limits of the parameters. The  constraint results are shown in 
   Figure \ref{fig:1}-\ref{fig:6} and summarized in Table \ref{tab:5}.

   \begin{figure}[htbp]
      \centering
      \subfigure{
         \includegraphics[width=0.22\textwidth]{ohd_1.pdf}
      }
      \subfigure{
         \includegraphics[width=0.22\textwidth]{ohd_2.pdf}
      }
      \caption{The constraint result from the observational
      $H(z)$ data.}
      \label{fig:1}
   \end{figure}

   \begin{figure}[htbp]
      \centering
      \subfigure{
         \includegraphics[width=0.22\textwidth]{sne_1.pdf}
      }
      \subfigure{
         \includegraphics[width=0.22\textwidth]{sne_2.pdf}
      }
      \caption{The constraint result from the supernovae
      Ia data.}
   \end{figure}

   \begin{figure}[htbp]
      \centering
      \subfigure{
         \includegraphics[width=0.22\textwidth]{sne_qso_1.pdf}
      }
      \subfigure{
         \includegraphics[width=0.22\textwidth]{sne_qso_2.pdf}
      }
      \caption{The constraint result from the SNe Ia + QSO data.}
   \end{figure}

   We could find that the combined result of SNe Ia and QSO is most
   similar to SNe Ia result, what may refer that the QSO cannot
   give an effective constraint. It may be due to the data quality
   at the high redshift, while the effective redshift of QSO as tracers in BAO
   is about 1.5, within the SNe Ia range.

   \begin{figure}[htbp]
      \centering
      \subfigure{
         \includegraphics[width=0.22\textwidth]{bao_1.pdf}
      }
      \subfigure{
         \includegraphics[width=0.22\textwidth]{bao_2.pdf}
      }
      \caption{The constraint result from the BAO data.}
   \end{figure}

   \begin{figure}[htbp]
      \centering
      \subfigure{
         \includegraphics[width=0.22\textwidth]{ohd_sne_1.pdf}
      }
      \subfigure{
         \includegraphics[width=0.22\textwidth]{ohd_sne_2.pdf}
      }
      \caption{The constraint result from the OHD + SNe Ia data.}
   \end{figure}

   \begin{figure}[htbp]
      \centering
      \subfigure{
         \includegraphics[width=0.22\textwidth]{ohd_sne_bao_1.pdf}
      }
      \subfigure{
         \includegraphics[width=0.22\textwidth]{ohd_sne_bao_2.pdf}
      }
      \caption{The constraint result from the OHD + SNe Ia + BAO data.}
      \label{fig:6}
   \end{figure}

   \begin{table*}[htbp]
      \centering
      \begin{tabular}{lcccccc}
         \hline\hline
         Model 1 & OHD & SNe Ia & SNe Ia + QSO & BAO &
          OHD + SNe Ia & OHD + SNe Ia + BAO \\
         \hline
         $\Omega_{2,0}$ & $0.321_{-0.067}^{+0.055}$ & $0.360_{-0.020}^{+0.020}$
          & $0.359_{-0.020}^{+0.020}$ & $0.301_{-0.013}^{+0.013}$
          & $0.330_{-0.016}^{+0.016}$ & $0.313_{-0.010}^{+0.010}$ \\
         $\log_{10}(\kappa C_1/$Gyr${}^{-1})$ & $<-2.17$ & $<-2.51$ & $<-2.57$
          & $<-3.10$ & $<-2.88$ & $<-3.39$ \\
         $\log_{10}(M_x/$GeV) & $>-9.77$ & $>-9.42$ & $>-9.40$
          & $>-8.84$ & $>-9.03$ & $>-8.52$ \\
         \hline
      \end{tabular}
      \caption{The constraint results of Model 1.}
      \label{tab:5}
   \end{table*}

@@

   \begin{figure}[htbp]
      \centering
      \subfigure{
         \includegraphics[width=0.22\textwidth]{ohd_widm_1.pdf}
      }
      \subfigure{
         \includegraphics[width=0.22\textwidth]{ohd_widm_2.pdf}
      }
      \caption{The constraint result from the observational
      $H(z)$ data.}
   \end{figure}

   \begin{figure}[htbp]
      \centering
      \subfigure{
         \includegraphics[width=0.22\textwidth]{sne_widm_1.pdf}
      }
      \subfigure{
         \includegraphics[width=0.22\textwidth]{sne_widm_2.pdf}
      }
      \caption{The constraint result from the supernovae
      Ia data.}
   \end{figure}

   \begin{figure}[htbp]
      \centering
      \subfigure{
         \includegraphics[width=0.22\textwidth]{sne_qso_widm_1.pdf}
      }
      \subfigure{
         \includegraphics[width=0.22\textwidth]{sne_qso_widm_2.pdf}
      }
      \caption{The constraint result from the SNe Ia + QSO data.}
   \end{figure}
